Kids PM Summary

### 2019-12-17 dustrak_data_cleaning.Rmd
general data cleaning

---
title: "KidsAIR initial PM data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 13 Dec 2019"
output: pdf_document
---

@@

# Load, format, and save KidsAIR Dusttrak data
```{r}
##### Read in dusttrack .txt files (from WMT) #####

@@

# Load and combine data sheets with QAQC comments, etc
ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv") %>% 
  mutate(area = "AK")
ak_dt <- read_xlsx("Input/AK/ak_dustrak.xlsx") %>% 
  mutate(area = "AK") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area)
nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv") %>% 
  mutate(area = "NN")
nn_dt <- read_xlsx("Input/NN/nn_dustrak.xlsx") %>% 
  mutate(area = "NN") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area)
wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv") %>% 
  mutate(area = "WMT")
wmt_dt <- read_xlsx("Input/WMT/wmt_dustrak.xlsx") %>% 
  mutate(area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area)

@@

##### Add to new variables to Kids PM data #####
kids_pm_new <- kids_pm_joined %>%  
  #filter(area == "AK") %>% 
  # apply correction factor to pm and change to ug
  mutate(pm = pm/1.65,
         pm = pm*1000) %>% 
  mutate(home_winter_id = as.character(home_winter_id),
         area = as.factor(area),
         pm_date = date,
         pm_time = time,
         pm = as.numeric(pm),
         day_of_week = weekdays(date)) %>% 
  unite("pm_datetime", date:time, sep = " ") %>% 
  #select(home_winter_id, area, pm_date, pm_time, pm_datetime, day_of_week, pm) %>% 
  arrange(area, home_winter_id, pm_date) %>% 
  group_by(area, home_winter_id) %>% 
  # new variables summarizing pm2.5 within each home_winter_id
  mutate(pm_total_observations = n(),
         # new datetime values to calculate sampling interval
         min_datetime = first(pm_datetime),
         max_datetime = last(pm_datetime),
         min_datetime = ymd_hms(min_datetime),
         max_datetime = ymd_hms(max_datetime),
         diff_datetime = interval(min_datetime, max_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = pm_sample_interval*24*60,
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         pm_mean = mean(pm),
         # breaking sampling period into days
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
  # calculate pm means for each sampling day
  group_by(area, home_winter_id, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm),
         pm_sampling_day = sampling_day) %>% 
  ungroup() %>% 
  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day, -datetime)

@@

>>>>

---
title: "KidsAIR initial PM data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 16 Dec 2019"
output: pdf_document
---

@@

# Initial load, format, and save KidsAIR Dusttrak data
```{r}
##### Read in dusttrack .txt files (from WMT) #####

@@

# Load and combine data sheets with QAQC comments, etc
ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv") %>% 
  mutate(area = "AK")
ak_dt <- read_xlsx("Input/AK/ak_dustrak.xlsx") %>% 
  mutate(area = "AK") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv") %>% 
  mutate(area = "NN")
nn_dt <- read_xlsx("Input/NN/nn_dustrak.xlsx") %>% 
  mutate(area = "NN") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)
wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv") %>% 
  mutate(area = "WMT")
wmt_dt <- read_xlsx("Input/WMT/wmt_dustrak.xlsx") %>% 
  mutate(area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, dtid, area, startdate, starttime, stopdate, stoptime)

@@

##### Add to new variables to Kids PM data #####
kids_pm_new <- kids_pm_joined %>%  
  #filter(area == "AK") %>% 
  # apply correction factor to pm and change to ug
  mutate(pm = pm/1.65,
         pm = pm*1000) %>% 
  mutate(home_winter_id = as.character(home_winter_id),
         area = as.factor(area),
         pm_date = date,
         pm_time = time,
         pm = as.numeric(pm),
         day_of_week = weekdays(date)) %>% 
  unite("pm_datetime", date:time, sep = " ") %>% 
  #select(home_winter_id, area, pm_date, pm_time, pm_datetime, day_of_week, pm) %>% 
  arrange(area, home_winter_id, pm_date) %>% 
  group_by(area, home_winter_id) %>% 
  # new variables summarizing pm2.5 within each home_winter_id
  mutate(pm_total_observations = n(),
         # new datetime values to calculate sampling interval
         min_datetime = first(pm_datetime),
         max_datetime = last(pm_datetime),
         min_datetime = ymd_hms(min_datetime),
         max_datetime = ymd_hms(max_datetime),
         diff_datetime = interval(min_datetime, max_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = pm_sample_interval*24*60,
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         pm_mean = mean(pm),
         # breaking sampling period into days
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
  # calculate pm means for each sampling day
  group_by(area, home_winter_id, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm),
         pm_sampling_day = sampling_day) %>% 
  separate(starttime, c("trash1", "starttime"), sep = " ") %>% 
  separate(stoptime, c("trash2", "stoptime"), sep = " ") %>% 
  ungroup() %>% 
  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day, 
         -datetime, -trash1, -trash2)

@@

# Summarize by ID and Dusttrak ID to look for trends
```{r}
# Load data
kids_pm <- read_rds("Output/kids_pm_new.rds")

# summary by home_winter_id
summary_id <- kids_pm %>% 
  group_by(area, home_winter_id) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Act vs Exp Diff" = first(as.numeric(actual_vs_expected_diff)),
            "Act vs Exp" = first(actual_vs_expected),
            "Comments" = first(qaqccomments))
 write_csv(summary_id, "output/kids_summary_id.csv")

# summary by dusttrak id
summary_dtid <- kids_pm %>% 
  group_by(dtid) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))
```

# Histogram for distribution of PM by individual ID
```{r}
# Load data
kids_pm <- read_rds("Output/kids_pm_new.rds")

histogram_id <- kids_pm %>% 
  filter(home_winter_id == "10" & area == "AK") %>% 
  ggplot() + 
    geom_histogram(aes(pm), bins = 50) +
    theme_classic()
histogram_id
```

# Function to look at PM data across time
```{r}
# Plot PM by datetime to look at trends, gaps, and peaks
pm_time_plot_function <- function(id, location) {
pm_time_plot <- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  #filter(pm < 30000) %>% 
  arrange(pm_datetime) %>% 
  mutate(pm_sampling_day = as.character(pm_sampling_day),
         day_of_week = factor(day_of_week,
                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
                                         "Thursday", "Friday", "Saturday"))) %>%
  ggplot() + 
    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
    theme_classic() +
    labs(title = id,
         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
         x = "Datetime",
         color = "Day of week") +
    theme(axis.title.y = element_text(size = 12,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_text(size = 12),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 12, color = "black"),
          axis.line.x = element_line(colour = "black", size = 1), 
          axis.line.y = element_line(colour = "black", size = 1), 
          axis.ticks = element_blank()) +
    scale_color_manual(values = jvPalette)

gaps <<- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  mutate(pm_datetime = ymd_hms(pm_datetime),
         datetime_lead = ymd_hms(lead(pm_datetime)),
         datetime_1 = seconds(pm_datetime),
         datetime_2 = seconds(datetime_lead),
         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
  arrange(desc(datetime_diff_minutes)) %>% 
  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
         pm_datetime, datetime_lead)

pm_desc <<- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  arrange(desc(pm)) %>% 
  select(pm_datetime, pm, pm_sampling_day)

pm_time_plot
}
```

# Run function from above to look at results
```{r}
# Load data
kids_pm <- read_rds("Output/kids_pm_new.rds")

# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "512", location = "WMT")
head(gaps, 25)
head(pm_desc, 25)
tail(pm_desc, 25)

# filter data for a single home
check_data <- kids_pm %>% 
  filter(area == "AK" & home_winter_id == "10")

%>% 
  filter(pm < 0)
```

# Check sample start times
```{r}
# Load data
kids_pm <- read_rds("Output/kids_pm_new.rds")

# Checking difference between dustrak start datetime and field data log datetime
sample_times <- kids_pm %>% 
  #filter(area == "AK") %>% 
  group_by(area, home_winter_id) %>% 
  arrange(pm_datetime) %>% 
  unite(start_datetime, c("startdate", "starttime"), 
        sep = " ", remove = FALSE) %>% 
  mutate(first_datetime = first(pm_datetime),
         first_datetime = ymd_hms(first_datetime),
         start_datetime = ymd_hms(start_datetime),
         time_diff = first_datetime - start_datetime) %>% 
  distinct(time_diff, .keep_all = TRUE) %>% 
  select(area, home_winter_id, pm_datetime, first_datetime,
         start_datetime, time_diff, startdate, starttime) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id)
#write_csv(sample_times, "Output/pm_sample_times_comparison.csv")
```

# Loading and combining notes from time and PM logs
```{r}
time_notes <- read_xlsx("Input/pm_sample_times_comparison.xlsx") %>% 
  select(area, home_winter_id, first_datetime, start_datetime,
         time_diff, next_steps_time, check_log_time)

pm_notes <- read_xlsx("Input/kids_summary_id.xlsx") %>% 
  select(area, home_winter_id, n, further_comments_pm, next_steps_pm, check_log_pm)

kids_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  mutate(home_winter_id = as.numeric(home_winter_id))

notes_joined <- pm_notes %>% 
  left_join(time_notes, by = c("area", "home_winter_id")) %>% 
  left_join(kids_ids, by = c("area", "home_winter_id"))
#write_csv(notes_joined, "Output/pm_notes_check_logs.csv")
```

### 2019-12-17 home_child_data_cleaning.Rmd
general data cleaning

---
title: "KidsAIR initial home/child data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 13 Dec 2019"
output: pdf_document
---

@@

# Load, format, and save home and child overview/demographic data
```{r}
# Load files
wmt_child_initial <- read_xlsx("Input/wmt_child_20191126.xlsx")
wmt_home_initial <- read_xlsx("Input/wmt_home_20191126.xlsx")
wmt_homewinter_initial <- read_xlsx("Input/wmt_homewinter_20191126.xlsx")
wmt_parent_demographics_initial <- read_xlsx("Input/wmt_parent_demographics_20191126.xlsx")

# join all files from above
joined1 <- wmt_child_initial %>% 
  mutate(hispanic_child = as.factor(Hispanic),
         race_child = as.factor(Race),
         comments_child = Comments,
         what_missing_child = WhatMissing,
         incomplete_child = Incomplete,
         ignore_missing_child = IgnoreMissingData,
         ignore_reason_child = IgnoreReason) %>% 
  select(-ID, -Hispanic, -Race, -Comments, -WhatMissing, -Incomplete,
         -IgnoreMissingData, -IgnoreReason) %>% 
  left_join(wmt_homewinter_initial, by = "HomeID") %>% 
  left_join(wmt_home_initial, by = "HomeID") %>% 
  left_join(wmt_parent_demographics_initial, by = "HomeWinterID") %>%
  mutate(gender_parent = as.factor(Gender),
         hispanic_parent = as.factor(Hispanic),
         race_parent = as.factor(Race),
         what_missing_parent = WhatMissing,
         incomplete_parent = Incomplete,
         ignore_missing_parent = IgnoreMissingData,
         ignore_reason_parent = IgnoreReason) %>%
  select(-ID, -Hispanic, -Race, -WhatMissing, -Incomplete,
         -IgnoreMissingData, -IgnoreReason, -Gender, -EnterDate) %>%
  rename_all(tolower)

# create dataset of id's and treatment assignment
wmt_ids_linked <- joined1 %>% 
  select(childid, homeid, home, homewinterid, winterid, txcondition) %>% 
  mutate(child_id = as.character(childid),
         home_id = as.character(homeid),
         home_winter_id = as.character(homewinterid),
         winter_id = as.character(winterid),
         treatment = as.factor(txcondition)) %>% 
  select(-childid, -homeid, -homewinterid, -winterid, -txcondition)

# create dataset of demographic data
wmt_demographic_data <- joined1 %>% 
  mutate(child_id = as.character(childid),
         home_id = as.character(homeid),
         age_nov1 = as.factor(agenov1),
         child_assent_date = ymd(childassentdate),
         parent_consent_date = ymd(parentconsentdate),
         home_winter_id = as.character(homewinterid),
         total_residents = as.numeric(totalres),
         total_residents_under5 = as.numeric(underfive),
         ignore_blanks_parent = ignoreblanks,
         income = as.factor(income),
         education = as.factor(education)) %>% 
  select(child_id, home_id, hispanic_child, age_nov1, race_child, child_assent_date,
         parent_consent_date, comments_child, what_missing_child, incomplete_child,
         enrolled, ignore_missing_child, ignore_reason_child,
         home_winter_id, gender_parent, hispanic_parent, race_parent, income,
         education, total_residents, total_residents_under5, incomplete_parent,
         ignore_blanks_parent, what_missing_parent, ignore_missing_parent,
         ignore_reason_parent) %>% 
  distinct(child_id, .keep_all = TRUE)

# create dataset of data on home details
## lower on priority list - see which variables we will use and clean later
wmt_home_details_uncleaned <- wmt_homewinter_initial %>%
  left_join(wmt_home_initial, by = "HomeID") %>% 
  rename_all(tolower)

# Save as RDS and CSV
write_rds(wmt_ids_linked, "Output/wmt_ids_linked.rds")
write_rds(wmt_demographic_data, "Output/wmt_demographic_data.rds")
write_csv(wmt_ids_linked, "Output/wmt_ids_linked.csv")
write_csv(wmt_demographic_data, "Output/wmt_demographic_data.csv")
```

```{r}

```

>>>>

---
title: "KidsAIR initial home/child data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 16 Dec 2019"
output: pdf_document
---

@@

# Load, format, and save home and child overview/demographic data for WMT
```{r}
# Load files
wmt_child_initial <- read_xlsx("Input/WMT/wmt_child_20191126.xlsx") %>% 
  rename_all(tolower) %>% 
  select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
  mutate(child_id_num = as.character(childid),
         home_id = as.character(homeid),
         child_id_char = as.character(id),
         hispanic_child = as.factor(hispanic),
         race_child = as.factor(race),
         age_child_nov = as.character(agenov1)) %>% 
  select(child_id_num, home_id, child_id_char, hispanic_child,
         race_child, age_child_nov, enrolled)

wmt_home_initial <- read_xlsx("Input/WMT/wmt_home_20191126.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homeid, home, txcondition) %>% 
  mutate(home_id = as.character(homeid),
         home = as.character(home),
         treatment = as.factor(txcondition)) %>% 
  select(home_id, home, treatment)

wmt_homewinter_initial <- read_xlsx("Input/WMT/wmt_homewinter_20191126.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         home_id = as.character(homeid),
         winter_id = as.character(winterid),
         intervention_date = interventiondate,
         intervention_time = interventiontime) %>% 
  select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)

wmt_parent_demographics_initial <-
  read_xlsx("Input/WMT/wmt_parent_demographics_20191126.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, gender, hispanic, race, income, education,
         totalres, underfive) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         gender_parent = as.factor(gender),
         hispanic_parent = as.factor(hispanic),
         race_parent = as.factor(race),
         income = as.factor(income),
         education = as.factor(education),
         total_residents = as.character(totalres),
         residents_under_five = as.character(underfive)) %>% 
  select(home_winter_id, gender_parent, hispanic_parent, race_parent,
         income, education, total_residents, residents_under_five) %>% 
  filter(!is.na(gender_parent))

# join all files from above
joined1 <- wmt_homewinter_initial %>% 
  left_join(wmt_parent_demographics_initial, by = "home_winter_id") %>% 
  left_join(wmt_home_initial, by = "home_id") %>% 
  left_join(wmt_child_initial, by = "home_id") %>% 
  mutate(area = "WMT")

# create dataset of id's and treatment assignment
wmt_home_ids_linked <- joined1 %>% 
  select(area, home_id, winter_id, home, home_winter_id, treatment) %>% 
  distinct(home_winter_id, .keep_all = TRUE) %>% 
  arrange(home_id, winter_id)

# create dataset of child demographic data
wmt_child_demographics <- joined1 %>% 
  distinct(child_id_num, .keep_all = TRUE) %>% 
  arrange(home_id, child_id_num, winter_id) %>% 
  filter(enrolled == 1)

# Save as RDS and CSV
write_rds(wmt_home_ids_linked, "Output/wmt_ids_linked.rds")
write_rds(wmt_child_demographics, "Output/wmt_demographic_data.rds")
write_csv(wmt_home_ids_linked, "Output/wmt_ids_linked.csv")
write_csv(wmt_child_demographics, "Output/wmt_demographic_data.csv")
```

# Load, format, and save home and child overview/demographic data for NN
```{r}
# Load files
nn_child_initial <- read_xlsx("Input/NN/nn_child_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
  mutate(child_id_num = as.character(childid),
         home_id = as.character(homeid),
         child_id_char = as.character(id),
         hispanic_child = as.factor(hispanic),
         race_child = as.factor(race),
         age_child_nov = as.character(agenov1)) %>% 
  select(child_id_num, home_id, child_id_char, hispanic_child,
         race_child, age_child_nov, enrolled)

nn_home_initial <- read_xlsx("Input/NN/nn_home_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homeid, home, txcondition) %>% 
  mutate(home_id = as.character(homeid),
         home = as.character(home),
         treatment = as.factor(txcondition)) %>% 
  select(home_id, home, treatment)

nn_homewinter_initial <- read_xlsx("Input/NN/nn_homewinter_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         home_id = as.character(homeid),
         winter_id = as.character(winterid),
         intervention_date = interventiondate,
         intervention_time = interventiontime) %>% 
  select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)

nn_parent_demographics_initial <-
  read_xlsx("Input/NN/nn_parent_demographics_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, gender, hispanic, race, income, education,
         totalres, underfive) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         gender_parent = as.factor(gender),
         hispanic_parent = as.factor(hispanic),
         race_parent = as.factor(race),
         income = as.factor(income),
         education = as.factor(education),
         total_residents = as.character(totalres),
         residents_under_five = as.character(underfive)) %>% 
  select(home_winter_id, gender_parent, hispanic_parent, race_parent,
         income, education, total_residents, residents_under_five) %>% 
  filter(!is.na(gender_parent))

# join all files from above
joined1 <- nn_homewinter_initial %>% 
  left_join(nn_parent_demographics_initial, by = "home_winter_id") %>% 
  left_join(nn_home_initial, by = "home_id") %>% 
  left_join(nn_child_initial, by = "home_id") %>% 
  mutate(area = "NN")

# create dataset of id's and treatment assignment
nn_home_ids_linked <- joined1 %>% 
  select(area, home_id, winter_id, home, home_winter_id, treatment) %>% 
  distinct(home_winter_id, .keep_all = TRUE) %>% 
  arrange(home_id, winter_id)

# create dataset of child demographic data
nn_child_demographics <- joined1 %>% 
  distinct(child_id_num, .keep_all = TRUE) %>% 
  arrange(home_id, child_id_num, winter_id) %>% 
  filter(enrolled == 1)

# Save as RDS and CSV
write_rds(nn_home_ids_linked, "Output/nn_ids_linked.rds")
write_rds(nn_child_demographics, "Output/nn_demographic_data.rds")
write_csv(nn_home_ids_linked, "Output/nn_ids_linked.csv")
write_csv(nn_child_demographics, "Output/nn_demographic_data.csv")
```

# Load, format, and save home and child overview/demographic data for AK
```{r}
# Load files
ak_child_initial <- read_xlsx("Input/AK/ak_child_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(childid, homeid, id, hispanic, race, agenov1, enrolled) %>% 
  mutate(child_id_num = as.character(childid),
         home_id = as.character(homeid),
         child_id_char = as.character(id),
         hispanic_child = as.factor(hispanic),
         race_child = as.factor(race),
         age_child_nov = as.character(agenov1)) %>% 
  select(child_id_num, home_id, child_id_char, hispanic_child,
         race_child, age_child_nov, enrolled)

ak_home_initial <- read_xlsx("Input/AK/ak_home_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homeid, home, txcondition) %>% 
  mutate(home_id = as.character(homeid),
         home = as.character(home),
         treatment = as.factor(txcondition)) %>% 
  select(home_id, home, treatment)

ak_homewinter_initial <- read_xlsx("Input/AK/ak_homewinter_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         home_id = as.character(homeid),
         winter_id = as.character(winterid),
         intervention_date = interventiondate,
         intervention_time = interventiontime) %>% 
  select(home_winter_id, home_id, winter_id, intervention_date, intervention_time)

ak_parent_demographics_initial <-
  read_xlsx("Input/AK/ak_parent_demographics_20191216.xlsx") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, gender, hispanic, race, income, education,
         totalres, underfive) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         gender_parent = as.factor(gender),
         hispanic_parent = as.factor(hispanic),
         race_parent = as.factor(race),
         income = as.factor(income),
         education = as.factor(education),
         total_residents = as.character(totalres),
         residents_under_five = as.character(underfive)) %>% 
  select(home_winter_id, gender_parent, hispanic_parent, race_parent,
         income, education, total_residents, residents_under_five) %>% 
  filter(!is.na(gender_parent))

# join all files from above
joined1 <- ak_homewinter_initial %>% 
  left_join(ak_parent_demographics_initial, by = "home_winter_id") %>% 
  left_join(ak_home_initial, by = "home_id") %>% 
  left_join(ak_child_initial, by = "home_id") %>% 
  mutate(area = "AK")

# create dataset of id's and treatment assignment
ak_home_ids_linked <- joined1 %>% 
  select(area, home_id, winter_id, home, home_winter_id, treatment) %>% 
  distinct(home_winter_id, .keep_all = TRUE) %>% 
  arrange(home_id, winter_id)

# create dataset of child demographic data
ak_child_demographics <- joined1 %>% 
  distinct(child_id_num, .keep_all = TRUE) %>% 
  arrange(home_id, child_id_num, winter_id) %>% 
  filter(enrolled == 1)

# Save as RDS and CSV
write_rds(ak_home_ids_linked, "Output/ak_ids_linked.rds")
write_rds(ak_child_demographics, "Output/ak_demographic_data.rds")
write_csv(ak_home_ids_linked, "Output/ak_ids_linked.csv")
write_csv(ak_child_demographics, "Output/ak_demographic_data.csv")
```

# Combine linked id data for WMT, NN, AK
```{r}
ak_ids <- read_rds("Output/ak_ids_linked.rds")
nn_ids <- read_rds("Output/nn_ids_linked.rds")
wmt_ids <- read_rds("Output/wmt_ids_linked.rds")

kids_ids <- rbind(ak_ids, nn_ids, wmt_ids)

#write_rds(kids_ids, "Output/kids_linked_ids.rds")
```

# Combine demographic data for WMT, NN, AK
```{r}
ak_demographics <- read_rds("Output/ak_demographic_data.rds")
nn_demographics <- read_rds("Output/nn_demographic_data.rds")
wmt_demographics <- read_rds("Output/wmt_demographic_data.rds")

kids_demographics <- rbind(ak_demographics, nn_demographics, wmt_demographics)

#write_rds(kids_demographics, "Output/kids_combined_demographics.rds")
```

### 2019-12-17 pm_summary.Rmd
general data cleaning

---
title: "Kids PM Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019, Updated 13 Dec 2019"
output: pdf_document
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
library(lubridate)
jvPalette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

@@

# Summarize by ID and Dusttrak ID
```{r}
# summary by home_winter_id
summary_id <- kids_pm %>% 
  group_by(area, home_winter_id) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Act vs Exp Diff" = first(as.numeric(actual_vs_expected_diff)),
            "Act vs Exp" = first(actual_vs_expected),
            "Comments" = first(qaqccomments))
# write_csv(summary_id, "output/kids_summary_id.csv")

# summary by dusttrak id
summary_dtid <- kids_pm %>% 
  group_by(dtid) %>% 
  summarise("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))
```

# Histogram for distribution of PM by individual ID
```{r}
histogram_id <- kids_pm %>% 
  filter(home_winter_id == "10" & area == "AK") %>% 
  ggplot() + 
    geom_histogram(aes(pm), bins = 50) +
    theme_classic()
histogram_id
```

# Function to look at PM data across time
```{r}
# Plot PM by datetime to look at trends, gaps, and peaks
pm_time_plot_function <- function(id, location) {
pm_time_plot <- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  #filter(pm < 30000) %>% 
  arrange(pm_datetime) %>% 
  mutate(pm_sampling_day = as.character(pm_sampling_day),
         day_of_week = factor(day_of_week,
                              levels = c("Sunday", "Monday", "Tuesday", "Wednesday", 
                                         "Thursday", "Friday", "Saturday"))) %>%
  ggplot() + 
    geom_point(aes(pm_datetime, pm, color = day_of_week), size = 1.5) +
    theme_classic() +
    labs(title = id,
         y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")")),
         x = "Datetime",
         color = "Day of week") +
    theme(axis.title.y = element_text(size = 12,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_text(size = 12),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 12, color = "black"),
          axis.line.x = element_line(colour = "black", size = 1), 
          axis.line.y = element_line(colour = "black", size = 1), 
          axis.ticks = element_blank()) +
    scale_color_manual(values = jvPalette)

gaps <<- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  mutate(pm_datetime = ymd_hms(pm_datetime),
         datetime_lead = ymd_hms(lead(pm_datetime)),
         datetime_1 = seconds(pm_datetime),
         datetime_2 = seconds(datetime_lead),
         datetime_diff_minutes = as.numeric(datetime_2 - datetime_1)/60,
         datetime_diff_hours = as.numeric(datetime_diff_minutes/60)) %>% 
  arrange(desc(datetime_diff_minutes)) %>% 
  select(day_of_week, pm_sampling_day, datetime_diff_minutes, datetime_diff_hours, 
         pm_datetime, datetime_lead)

pm_desc <<- kids_pm %>% 
  filter(home_winter_id == id & area == location) %>% 
  arrange(desc(pm)) %>% 
  select(pm_datetime, pm, pm_sampling_day)

pm_time_plot
}
```

# Run function from above to look at results
```{r}
# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "85", location = "AK")
head(gaps, 25)
head(pm_desc, 25)
tail(pm_desc, 25)

# filter data for a single home
check_data <- kids_pm %>% 
  filter(area == "WMT" & home_winter_id == "513") %>% 
  filter(pm < 0)
```

>>>>

---
title: "Kids PM Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019, Updated 16 Dec 2019"
output: pdf_document
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(readxl)
library(knitr)
library(lubridate)
jvPalette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

@@
